---
title: "Understanding a Basic Shiny Application"
author: "Data Analytics Team"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

## Introduction
This document explains the structural architecture and components of a basic **R Shiny** web application. Shiny applications allow data analysts to turn R code into interactive web pages without needing to write HTML, CSS, or JavaScript.

---

## 1. The Live Interactive Application
Below is the running application. You can move the slider to see how the histogram dynamically changes its bins in real-time.

```{r application-demo, echo=FALSE}
# 1. User Interface Layout
ui_demo <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      sliderInput(
        inputId = "bins",
        label = "Number of bins:",
        min = 1,
        max = 50,
        value = 30
      )
    ),
    mainPanel(
      plotOutput(outputId = "distPlot")
    )
  )
)

# 2. Server Processing Logic
server_demo <- function(input, output) {
  output$distPlot <- renderPlot({
    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)
    
    hist(
      x, 
      breaks = bins, 
      col = "#75AADB", 
      border = "white",
      xlab = "Waiting time to next eruption (in mins)",
      main = "Histogram of Waiting Times"
    )
  })
}

# Run the app locally within the document
shinyApp(ui = ui_demo, server = server_demo, options = list(height = 400))
```

---

## 2. Core Architecture
A standard Shiny application relies on a **Client-Server** architecture split into two core script elements:

1. **The User Interface (UI):** The frontend web page that controls visual formatting, text headers, and input controls.
2. **The Server:** The backend execution system that processes calculations, filters data structures, and renders graphics using R.

Communication flows between these elements instantly via **Reactivity**. When an input value changes on the UI layout, the backend server updates its specific dependencies automatically.

---

## 3. Code Structural Breakdown

### The Frontend (UI Elements)
```r
ui <- fluidPage(
  titlePanel("My First Shiny App"),
  sidebarLayout(
    sidebarPanel(
      sliderInput(inputId = "bins", label = "Number of bins:", min = 1, max = 50, value = 30)
    ),
    mainPanel(
      plotOutput(outputId = "distPlot")
    )
  )
)
```
* **`fluidPage()`**: Sets up a responsive grid container that stretches or shrinks with the user's browser window size.
* **`sidebarLayout()`**: Formats the page layout into a 1/3-width left-hand input sidebar and a 2/3-width right-hand main panel.
* **`sliderInput()`**: Generates an HTML slider bar. The variable `inputId = "bins"` establishes the backend bridge variable.
* **`plotOutput()`**: reserves a visual grid coordinate on the front page named `"distPlot"` to capture graphics rendered by the backend.

### The Backend (Server Elements)
```r
server <- function(input, output) {
  output\$distPlot <- renderPlot({
    x    <- faithful\$waiting
    bins <- seq(min(x), max(x), length.out = input\$bins + 1)
    hist(x, breaks = bins, col = "#75AADB", border = "white")
  })
}
```
* **`input` & `output`**: Core list arguments processing incoming client data (`input$bins`) and returning calculated graphic parameters (`output$distPlot`).
* **`renderPlot()`**: A reactive calculation wrapper. It sets up an active listener. If `input$bins` changes, the script block inside the curly brackets `{}` triggers a complete visual recalculation.
